Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-10-31 thirdPaper/draft/typeConeAssociahedra.tex
FPSAC

\begin{theorem}
\label{thm:associahedronLoday}
The associahedron~$\Asso[n]$ can be described equivalently as:
\begin{itemize}
\item the convex hull of the points~$L(T) \in \R^{n+1}$ for all rooted binary trees~$T$ with~$n+1$ internal nodes, where the $i$th coordinate of~$L(T)$ is the product of the number of leaves in the left subtree by the number of leaves in the right subtree of the $i$th node of~$T$ in inorder~\cite{Loday},
\item the intersection of the hyperplane~$\bigset{\b{x} \in \R^{n+1}}{\sum_{\ell \in [n+1]} x_\ell = \binom{n+2}{2}}$ with the halfspaces $\bigset{\b{x} \in \R^{n+1}}{\sum_{a < \ell < b} x_\ell \ge \binom{b-a+2}{2}}$ for all internal diagonals~$(a,b)$ of \mbox{the $(n+3)$-gon~\cite{ShniderSternberg}},
\item the Minkowski sum of the faces~$\simplex_{[i,j]}$ of the standard $n$-dimensional simplex~$\simplex_{[n+1]}$ corresponding to all proper intervals~$[i,j]$ of~$[n+1]$,~\cite{Postnikov}.
\end{itemize}
\end{theorem}

@@

\begin{theorem}
\label{thm:normalFanLoday}
In the normal fan~$\gvectorFan[n]$ of~$\Asso[n]$, 
\begin{itemize}
\item the normal vector of the facet corresponding to an internal diagonal~$(a,b)$ of the $(n+3)$-gon is the vector~$\gvector{a,b} \eqdef \pi\big( \sum_{a < \ell < b} \b{e}_\ell \big) = (n+1) \sum_{a < \ell < b} \b{e}_\ell - (b-a+1) \sum_{1 \le \ell \le n+1} \b{e}_\ell$.
\item the normal cone of the vertex corresponding to a rooted binary tree~$T$ is the incidence cone~$\bigset{\b{x} \in \HH}{x_i \le x_j \text{ for all edges $i \to j$ in $T$}}$.
\end{itemize}
\end{theorem}

@@

The type cone of~$\gvectorFan[2]$ lives in~$\R^9$ and has a linearity space of dimension~$3$.
It has $6$ facet-defining inequalities (given below), which correspond to the flips described in \cref{prop:extremalExchangeablePairsAsso} and illustrated in \cref{fig:labelFacetDefiningInequalititiesAsso}\,(right).

@@

\enlargethispage{.2cm}
Combining \cref{coro:simplicialTypeCone}, \cref{coro:simplicialTypeConeAsso} and \cref{prop:extremalExchangeablePairsAsso}, we derive the following description of all polytopal realizations of the fan~$\gvectorFan[n]$, recovering all associahedra of~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan}.
Note that the arguments of~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} were quite different from the present~approach.

@@

The \defn{blossoming quiver}~$\quiver\blossom$ of a gentle quiver is the gentle quiver obtained by completing all vertices of~$\quiver$ with additional incoming or outgoing \defn{blossoms} such that all vertices of~$\quiver$ become $4$-valent.
For instance, \cref{fig:exmNonkissing},(left) shows a blossoming quiver: the vertices of~$\quiver$ appear in black, while the blossom vertices of~$\quiver\blossom$ appear in white.
We now always assume that~$\quiver$ is a gentle quiver with blossoming quiver~$\quiver\blossom$.

@@

A \defn{peak} (resp.~\defn{deep}) of a walk~$\omega$ is a substring of~$\omega$ of length zero which is on top (resp.~at the bottom of~$\omega$).
A walk~$\omega$ is \defn{straight} if it has no peak or deep (\ie if~$\omega$ or~$\omega^{-1}$ is a path in~$\quiver\blossom$), and \defn{bending} otherwise.
We denote by~$\peaks{\omega}$ (resp.~$\deeps{\omega}$) the multisets of vertices of peaks (resp.~deeps) of~$\omega$.

@@

\item For any non-kissing facets~$F, F'$ of~$\NKC$ with~$F \ssm \{\omega\} = f' \ssm \{\omega'\}$, the distinguished substrings of~$\omega$ in~$F$ and of~$\omega'$ in~$F'$ coincide to a string~$\sigma$. Moreover, if we decompose~${\omega = \rho \sigma \tau}$ and~${\omega' = \rho' \sigma \tau'}$, then the facets~$F$ and~$F'$ both contain the walks~$\mu \eqdef \rho' \sigma \tau$ and~$\nu \eqdef \rho \sigma \tau'$.
\item The substring~$\sigma$ and thus the walks~$\mu$ and~$\nu$ only depend on the exchangeable walks~$\omega$ and~$\omega'$, and not on the adjacent non-kissing facet~$F $ and~$F'$.
\item The linear dependence between the $\b{g}$-vectors of~$F \cup F'$ is given by
\[
\gvector{\omega} + \gvector{\omega'} = \gvector{\mu} + \gvector{\nu}.
\]
\item The non-kissing fan of~$\quiver$ has the unique exchange relation property.
\item The $\b{c}$-vector orthogonal to all $\b{g}$-vectors~$\gvector{\lambda}$ for~$\lambda \in F \cap F'$ is the multiplicity vector~$\multiplicityVector_{\sigma}$ of the vertices of the substring~$\sigma$ of~$\omega$ and~$\omega'$.
\end{enumerate}

>>>>

\begin{theorem}
\label{thm:associahedronLoday}
The associahedron~$\Asso[n]$ can be described equivalently as:
\begin{itemize}
\item the convex hull of the points~$L(T) \in \R^{n+1}$ for all rooted binary trees~$T$ with~$n+1$ internal nodes, where the $i$th coordinate of~$L(T)$ is the product of the number of leaves in the left subtree by the number of leaves in the right subtree of the $i$th node of~$T$ in inorder~\cite{Loday},
\item the intersection of the hyperplane~$\bigset{\b{x} \in \R^{n+1}}{\sum_{\ell \in [n+1]} x_\ell = \binom{n+2}{2}}$ with the halfspaces $\bigset{\b{x} \in \R^{n+1}}{\sum_{a < \ell < b} x_\ell \ge \binom{b-a}{2}}$ for all internal diagonals~$(a,b)$ of \mbox{the $(n+3)$-gon~\cite{ShniderSternberg}},
\item the Minkowski sum of the faces~$\simplex_{[i,j]}$ of the standard $n$-dimensional simplex~$\simplex_{[n+1]}$ corresponding to all proper intervals~$[i,j]$ of~$[n+1]$,~\cite{Postnikov}.
\end{itemize}
\end{theorem}

@@

\begin{theorem}
\label{thm:normalFanLoday}
In the normal fan~$\gvectorFan[n]$ of~$\Asso[n]$, 
\begin{itemize}
\item the normal vector of the facet corresponding to an internal diagonal~$(a,b)$ of the $(n+3)$-gon is the vector~$\gvector{a,b} \eqdef \pi\big( \sum_{a < \ell < b} \b{e}_\ell \big) = (n+1) \sum_{a < \ell < b} \b{e}_\ell - (b-a-1) \sum_{1 \le \ell \le n+1} \b{e}_\ell$.
\item the normal cone of the vertex corresponding to a rooted binary tree~$T$ is the incidence cone~$\bigset{\b{x} \in \HH}{x_i \le x_j \text{ for all edges $i \to j$ in $T$}}$.
\end{itemize}
\end{theorem}

@@

The type cone of~$\gvectorFan[3]$ lives in~$\R^9$ and has a linearity space of dimension~$3$.
It has $6$ facet-defining inequalities (given below), which correspond to the flips described in \cref{prop:extremalExchangeablePairsAsso} and illustrated in \cref{fig:labelFacetDefiningInequalititiesAsso}\,(right).

@@

\enlargethispage{.2cm}
Combining \cref{coro:simplicialTypeCone,coro:simplicialTypeConeAsso} and \cref{prop:extremalExchangeablePairsAsso}, we derive the following description of all polytopal realizations of the fan~$\gvectorFan[n]$, recovering all associahedra of~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan}.
Note that the arguments of~\cite[Sect.~3.2]{ArkaniHamedBaiHeYan} were quite different from the present~approach.

@@

The \defn{blossoming quiver}~$\quiver\blossom$ of a gentle quiver~$\quiver$ is the gentle quiver obtained by completing all vertices of~$\quiver$ with additional incoming or outgoing \defn{blossoms} such that all vertices of~$\quiver$ become $4$-valent.
For instance, \cref{fig:exmNonkissing},(left) shows a blossoming quiver: the vertices of~$\quiver$ appear in black, while the blossom vertices of~$\quiver\blossom$ appear in white.
We now always assume that~$\quiver$ is a gentle quiver with blossoming quiver~$\quiver\blossom$.

@@

A \defn{peak} (resp.~\defn{deep}) of a walk~$\omega$ is a substring of~$\omega$ of length zero which is on top (resp.~at the bottom) of~$\omega$.
A walk~$\omega$ is \defn{straight} if it has no peak or deep (\ie if~$\omega$ or~$\omega^{-1}$ is a path in~$\quiver\blossom$), and \defn{bending} otherwise.
We denote by~$\peaks{\omega}$ (resp.~$\deeps{\omega}$) the multisets of vertices of peaks (resp.~deeps) of~$\omega$.

@@

\item For any non-kissing facets~$F, F'$ of~$\NKC$ with~$F \ssm \{\omega\} = F' \ssm \{\omega'\}$, the distinguished substrings of~$\omega$ in~$F$ and of~$\omega'$ in~$F'$ coincide to a string~$\sigma$. Moreover, if we decompose~${\omega = \rho \sigma \tau}$ and~${\omega' = \rho' \sigma \tau'}$, then the facets~$F$ and~$F'$ both contain the walks~$\mu \eqdef \rho' \sigma \tau$ and~$\nu \eqdef \rho \sigma \tau'$.
\item The substring~$\sigma$ and thus the walks~$\mu$ and~$\nu$ only depend on the exchangeable walks~$\omega$ and~$\omega'$, and not on the adjacent non-kissing facet~$F $ and~$F'$.
\item The linear dependence between the $\b{g}$-vectors of~$F \cup F'$ is given by
\[
\gvector{\omega} + \gvector{\omega'} = \gvector{\mu} + \gvector{\nu}.
\]
\item The non-kissing fan of~$\quiver$ has the unique exchange relation property.
\item The $\b{c}$-vector orthogonal to all $\b{g}$-vectors~$\gvector{\lambda}$ for~$\lambda \in F \cap F'$ is the multiplicity vector~$\multiplicityVector_{\sigma}$ of the vertices of the substring~$\sigma$ of~$\omega$ and~$\omega'$.
\end{enumerate}
